---
title: "Seguro Popular"
author: "Var Kurapati"
format: html
execute:
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(gt)
```

```{r}
#| cache: true
x <- sps |>
  select(treatment, t2_health_exp_3m, age, sex, education) |>
  drop_na() |>
  mutate(treatment = factor(if_else(treatment == 1, "Enrolled", "Not enrolled"),
                            levels = c("Not enrolled", "Enrolled")))
```

```{r}
x |>
  filter(t2_health_exp_3m <= 5000) |>
  ggplot(aes(x = treatment, y = t2_health_exp_3m, color = treatment)) +
  geom_jitter(width = 0.2, alpha = 0.4) +
  stat_summary(fun = mean, geom = "point", size = 4, shape = 18, color = "black") +
  labs(
    title = "Three-Month Out-of-Pocket Health Spending by Seguro Popular Enrollment",
    subtitle = "Black diamonds show group means; spending is capped at 5,000 pesos for visibility",
    x = "Treatment Group",
    y = "Out-of-Pocket Health Spending (Mexican Pesos)",
    color = "Treatment",
    caption = "Source: King et al. (2009) via primer.data::sps"
  )
```

```{r}
#| cache: true
fit_sp <- linear_reg(engine = "lm") |> fit(t2_health_exp_3m ~ treatment, data = x)
```

```{r}
tidy(fit_sp, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(c(estimate, conf.low, conf.high), \(x) signif(x, 3))) |>
  gt() |>
  tab_header(title = "Estimates and 95% Confidence Intervals for Parameters",
             subtitle = "Source: King et al. (2009) via primer.data::sps")
```

Public health insurance programs aim to reduce the financial burden of medical care on households, particularly out-of-pocket spending that can push families toward poverty. Using data from a randomized rollout of Mexico's Seguro Popular program, evaluated by King et al. (2009) across roughly 27,500 households in seven states, we estimate the average causal effect of enrollment on three-month out-of-pocket health spending. However, because t2_health_exp_3m is heavily right-skewed, with a median of 150 pesos but a maximum past 37,000, a linear model assuming normally distributed errors may poorly capture the true distribution of spending and could be unduly influenced by extreme outliers. We model t2_health_exp_3m, three-month out-of-pocket health spending in pesos, as a linear function of treatment, the randomized assignment to early Seguro Popular enrollment.







